---
title: "__Non-_Saccharomyces_ Validation Study at University of California Davis__"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---  
***
# Prepared by Daniel A. Dycus for LAFFORT
# 50 g/hL of Egide versus 60 mg/L $\mathrm{SO_2}$  
***  

```{r Prepare Libraries Import Data, include=FALSE}
if(!require(FSA)){install.packages("FSA")}  
if(!require(ggplot2)){install.packages("ggplot2")} 
if(!require(car)){install.packages("car")} 
if(!require(multcompView)){install.packages("multcompView")}
if(!require(lsmeans)){install.packages("lsmeans")} 
if(!require(grid)){install.packages("grid")} 
if(!require(nlme)){install.packages("nlme")}   
if(!require(lme4)){install.packages("lme4")} 
if(!require(lmerTest)){install.packages("lmerTest")}   
if(!require(rcompanion)){install.packages("rcompanion")}   

# Load libraries
library(rcompanion)
library(ggplot2)
library(FSA)
library(car)
library(multcompView)
library(lsmeans)
library(grid)
library(nlme)
library(lme4)

#Import Datasets

Data=read.csv(file = "c:/users/ddycus/OneDrive - LAFFORT/Documents/R/UCD_Egide/ucd_data_raw1.csv", header = TRUE, fileEncoding = "UTF-8-BOM")
SacData=read.csv(file = "c:/users/ddycus/OneDrive - LAFFORT/Documents/R/UCD_Egide/sacbenefit/sacbenefit.csv", header = TRUE, fileEncoding = "UTF-8-BOM")

```
# AA = Acetic Acid Bacteria  
# HU = *Hanseniaspora uvarum*  
# LB = *Lactobacillus brevis/hilgardii/fermentum*  
# LK = *Lactobacillus kunkeei*  
# LP = *Lactobacillus plantarum/casei/mali*  
# PDO = *Pediococcus*   
# PM = *Pichia membranifaciens/fermentans*  
# ZYGO = *Zygosaccharomyces*  

```{r}
#Stuff for Peer Review
#tiff("figure4New.tiff", units="in", width=7.25, height=6.5, res=600)

#Create dataframe, check header, get data summary
Data$Treatment = factor(Data$Treatment, levels = unique(Data$Treatment))
Data$Organism = factor(Data$Organism, levels = unique(Data$Organism))
headtail(Data)
str(Data)
summary(Data)
Summarize(Cells ~ Treatment + Organism, data=Data, digits = 3)

#Plot the mean value of Organisms vs. Treatment
total=Summarize(Cells ~ Treatment + Organism, data=Data, digits = 3)
a<-ggplot(data=total, aes(x=Treatment, y=mean, fill=Treatment))+ #create plot, use mean value for y
  labs(y="cells/mL")+ #label axis
  geom_bar(stat="identity")+ #geom_bar
  theme_minimal()+ #minimal theme
  ggtitle("Average Amount of Spoilage Organisms")+ #title
  #Center the Title, change for guidelines per publication
  theme(plot.title = element_text(hjust = 0.5, color = "black", size= 14))
 
a + scale_fill_manual(values=c("#457b9d","#1d3557"))

# Stuff for Peer Review
# dev.off() 

```

# __Build Model__
```{r}
#Libraries
library(multcompView)
library(lsmeans)

# ANOVA Type II
model=lm(Cells ~ Treatment + Organism + Treatment:Organism, data=Data)
Anova(model, type = "II")

# Least Squares
marginal = lsmeans(model,
                   ~ Treatment)
pairs(marginal,
      adjust="tukey")

#Treatment Significance
CLD = cld.emmGrid(marginal,
          alpha   = 0.1,
          conf    = 0.95,
          Letters = letters,        
          adjust  = "tukey")  
CLD

#Letter Display for Organism, Note Hanseniaspora outlier, consider Monte Carlo, pub wants data raw.
marginal = lsmeans(model,
                   ~ Organism)
pairs(marginal,
      adjust="tukey")
CLD = cld.emmGrid(marginal,
          alpha   = 0.05,
          Letters = letters,     
          adjust  = "tukey")      
CLD

```

# __Plot of Least Square Means Comparing Treatments__
```{r}
#Use CLD to illustrate difference between the treatments
marginal = lsmeans(model,
                   ~ Treatment)
CLD = cld.emmGrid(marginal,
          alpha   = 0.1,
          conf    = 0.95,
          Letters = letters,        
          adjust  = "tukey")
CLD$treatment = factor(CLD$Treatment,
                        levels=c("Egide",
                                 "Td/Mp"))
CLD$.group=gsub(" ", "", CLD$.group)

library(ggplot2)
#Publication Requests
#tiff("figure1old.tiff", units="in", width=7.25, height=6.5, res=600)
ggplot(CLD, aes(x= Treatment, y = lsmean,label = .group)) +
    geom_point(shape  = 15,
               size   = 4) +
    geom_errorbar(aes(ymin  =  lsmean+SE, 
                      ymax  =  lsmean-SE), 
                      width =  0.2,
                      size  =  0.7) +

    theme_bw() +
    theme(plot.caption = element_text(hjust = 0.5),
         plot.title = element_text("Plot Name", hjust = 0.5),
         text = element_text(family="sans"), #Issue with font family, Editor and peer review doesn't want this graph. Making alternate.
         panel.background = element_rect(fill = 'white', color = 'white')) +

    ylab("Least Squares Mean\nWine Spoilage Organisms (cells / mL)") +
    ggtitle("Least squares means of organisms in response to treatment")
    geom_text(nudge_x = c(-.1,-.1),
            nudge_y = c(-0.1,-0.1),
            color   = "black")
 # Pub stuff
 # dev.off()
```
# __Plot for Organisms, Note Hanseniaspora__
```{r}
# Use Least Squares Means for Oranism. This information not making peer review. LSM used for internal validation only. Note Hanseniaspora 
marginal = lsmeans(model,
                   ~ Organism)
CLD = cld.emmGrid(marginal,
          alpha   = 0.1,
          conf    = 0.95,
          Letters = letters,        
          adjust  = "tukey")
CLD$Organism = factor(CLD$Organism,
                        levels=c("AA",
                                 "HU",
                                 "LB",
                                 "LK",
                                 "LP",
                                 "PDO",
                                 "PM",
                                 "ZYGO"))
CLD$.group=gsub(" ", "", CLD$.group)

ggplot(CLD,
       aes(x     = Organism,
           y     = lsmean,
           label = .group)) +

    geom_point(shape  = 15,
               size   = 4) +

    geom_errorbar(aes(ymin  =  lsmean+SE, 
                      ymax  =  lsmean-SE), 
                      width =  0.2,
                      size  =  0.7) +

    theme_bw() +
    theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +

    ylab("Least Squares Mean\nWine Spoilage Organisms (cells / mL)") +

  geom_text(nudge_x = c(.3,.3),
            nudge_y = c(-3,-0.3,-0.3,-0.3,-0.3,-0.3,-0.3,-0.3),
            color   = "black")

```
# __Means for Organisms and Treatments with Grouping Comparing Treatments__
```{r}

marginal = lsmeans(model,
                   ~ Treatment + Organism)
CLD = cld.emmGrid(marginal,
          alpha   = 0.1,
          Letters = letters,        
          adjust  = "tukey")      
### Order the levels for printing
CLD$Treatment = factor(CLD$Treatment,
                        levels=c("SO2",
                                 "Td/Mp"))
CLD$Organism = factor(CLD$Organism,
                        levels=c("AA",
                                 "HU",
                                 "LB",
                                 "LK",
                                 "LP",
                                 "PDO",
                                 "PM",
                                 "ZYGO"))
CLD
```
# __Least Square Mean Plot for Wine Spoilage Organisms, HU interfering with baseline__
```{r}
# Build Plot, this did not make peer review 
pd = position_dodge(0.6)    ### How much to jitter the points on the plot
ggplot(CLD,
       aes(x     = Treatment,
           y     = lsmean,
           color = Organism,
           label = .group)) +
    geom_point(shape  = 15,
               size   = 4,
             position = pd) +
    geom_errorbar(aes(ymin  =  lsmean+SE, #per Lucy Joseph
                      ymax  =  lsmean-SE), #per Lucy Joseph
                      width =  0.2,
                      size  =  0.7,
                     position = pd) +
    coord_cartesian(ylim=c(-1e6,1e7))+ #changing the boundaries
    theme_bw() +
    theme(axis.title   = element_text(face = "bold"),
          axis.text    = element_text(face = "bold"),
          plot.caption = element_text(hjust = 0)) +
  ylab("Least square mean\nWine Spoilage Organisms (cells / mL)") +
 
 # geom_text(nudge_x = c(-0.22, 0.08, -0.22, 0.22,
  #                       0.08, 0.08, -0.08, -0.22,
   #                      0.22, 0.22, -0.08, -0.08),
    #        nudge_y = rep(270, 12),
     #       color   = "black") +
 
  scale_color_manual(values = c("magenta", "palevioletred", "skyblue", "lawngreen", "turquoise3","orchid1","deeppink1","maroon1" ))
```

# __Saccharomyces Log Change__
```{r}

#Set up Data Frame
 df <- data.frame(Stage=rep(c("Cold Soak","Early","Mid","Late"), each=2),
        Treatment=rep(c("SO2","Td/Mp"),2),
        LogChange=c(3.29,5.42,1.32,1.58,-.60,-.85,-2.31,-2.38))
head(df)
# Set Fermentation Stage as a Factor
df$Stage <- factor(df$Stage, levels=c("Cold Soak", "Early", "Mid", "Late"))
# Pub Stuff
# tiff("figure2.tiff", units="in", width=7.25, height=6.5, res=600)
# Build the figure
ggplot(data=df, aes(x=Stage, y=LogChange, fill=Treatment))+
  ggtitle(~italic("S. cerevisae")~ "Log Change")+
  theme(plot.title = element_text(hjust = 0.5),
  plot.subtitle = element_text(hjust = 0.5),
  text=element_text(family = "sans"),
  panel.background = element_rect(fill='white', color = 'white'))+
  geom_bar(stat="identity", position=position_dodge())+
  scale_fill_manual(values = c("#457b9d","#1d3557"))
 # Pub Stuff
 # dev.off()
```
# __From Editor: Present Data Simply__
```{r}
#convert Day Column from numberic to factor variable
Data$Day <- as.factor(Data$Day)
head(Data)
# Pub Stuff
# tiff("figure1.tiff", units="in", width=7.25, height=6.5, res=600)
p <- ggplot(Data, aes(x=Day, y=Cells, fill=Treatment))+ 
  geom_col(linetype=0,color="black", position = position_dodge())+
  scale_fill_manual(values = c("#457b9d","#1d3557"))+
  ggtitle("Day of fermentation (cells/mL)")+
  theme(plot.title = element_text(hjust = 0.5, color = "black", size = 12), 
        text=element_text(family="sans"),
        panel.background = element_rect(fill='white', color = 'white'))
  
  p + scale_x_discrete(labels=unique(Data$Day))+
   labs(x = "Day of fermentation", y = "cells/mL")
 # Pub Stuff
 # dev.off()  

```

# __Seeking best visual representation of data__
```{r}
#Make a new figure for readers that is easier to read

#Save the file in 600 DPI for Catalyst
#tiff("figure6bad.tiff", units="in", width=7.25, height=6.5, res=600)

#Read CSV and create df
df <- read.csv(file = "c:/users/ddycus/OneDrive - LAFFORT/Documents/R/UCD_Egide/dammitjanet.csv", header = TRUE, fileEncoding = "UTF-8-BOM")

#Stack it up
p<-ggplot(Data, aes(x=Treatment, y=Cells, fill=Organism))+
  geom_bar(position="stack", stat="identity")
p  + scale_y_continuous(trans = 'log2')

# Remove the tiff for Catalyst
# dev.off()  

```
# __Additional Data Representation__
```{r}
#Make a new figure for Catalyst that is easier to read, stacked up don't look good, or change scale of the cells, log scale has worked before. Also was ineffective, shoot for one image with facets and main image.
library(gridExtra)

dammitjanet <- read.csv(file = "c:/users/ddycus/OneDrive - LAFFORT/Documents/R/UCD_Egide/dammitjanet.csv", header = TRUE, fileEncoding = "UTF-8-BOM")

df <- data.frame(Organism=unique(dammitjanet$Organism), Cells=(dammitjanet$Cells))

p1 <-ggplot(data=dammitjanet, aes(x=Treatment, y=Cells, fill=Treatment))+
  geom_bar(stat="identity")+ #type of plot
  labs(y="Sum of cells/mL")+ #y labels
  theme_minimal()+
  theme(legend.position = 'none')
  ggtitle("Sum of Spoilage Organisms")+
  #Center the Title, change for guidelines per ASEV
  theme(plot.title = element_text(hjust = 0.5, color = "black", size= 14))
p1 = p1 + scale_fill_manual(values=c("#457b9d","#1d3557"))
p1


```
# __A faceted approach__
```{r}
library(stringr)
library(plyr)
library(dplyr)
#Save the file in 600 DPI for Catalyst
#tiff("figure7bad.tiff", units="in", width=7.25, height=6.5, res=600)

#Wrap long labels
var_width = 10
dammitjanet1 <- mutate(dammitjanet, Organism = str_wrap(Organism, width = var_width))

#Making second plot with facets
p2 <- ggplot(dammitjanet1, aes(x=Treatment, y=Cells, factor(Organism), fill=Treatment))+
  geom_bar(stat = "identity") + facet_wrap(~Organism, ncol = 2, scales = "free")+
  theme(legend.position = "none")
p2 = p2 + theme(axis.title.x = element_blank())+
  scale_fill_manual(values=c("#457b9d","#1d3557"))

grid.arrange(p1,p2, widths = c(1.5,2))
# Remove the tiff for Catalyst
# dev.off()
```
# __Specific Request__
```{r}

# Specific request for image
# Save the file in 600 DPI for Catalyst
# tiff("figure9bad.tiff", units="in", width=7.25, height=6.5, res=600)

#Make a function to extrapolate the legend
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

#make the first one
df1 <- data.frame(Organism=rep(c("Total Cells","Hanseniaspora uvarum"), each=2),
        Treatment=rep(c("SO2","Td/Mp"),2),
        Cells=c(29012585, 9880348, 28309780, 9387320))

#make the second one
df2 <- data.frame(Organism=rep(c("Acetic Acid Bacteria","Lactobacillus kunkeei"), each=2),
                  Treatment=rep(c("SO2","Td/Mp"),2),
                  Cells=c(305040, 260620, 388900, 214650))

#make the third one
df3 <- data.frame(Organism=rep(c("Pediococcus Species","Zygosaccharomyces","L. brevis/hilgardii/fermentum","Lactobacillus plantarum/casei/mali", "Pichia membranifaciens/fermentans"), each=2),
                  Treatment=rep(c("SO2","Td/Mp"),5),
                  Cells=c(5450, 6290, 1451, 1291, 1101, 7763, 571, 1931, 292, 483))

#Wrap the long labels for each figure
var_width = 15
df1 <- mutate(df1, Organism = str_wrap(Organism, width = var_width))
df2 <- mutate(df2, Organism = str_wrap(Organism, width = var_width))
df3 <- mutate(df3, Organism = str_wrap(Organism, width = var_width))

# Build the first figure
p1 <- ggplot(data=df1, aes(x= reorder(Organism, -Cells), y=Cells, fill=Treatment))+
 # ggtitle("Total cells and"~italic("Hanseniaspora uvarum"))+
  theme(axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    text=element_text(family = "sans"),
    panel.background = element_rect(fill='white', color = 'white'))+
  geom_bar(stat="identity", position=position_dodge())+
  labs(tag = 'A')+ #add letters to fulfill multipanel requirements from Catalyst
  scale_fill_manual(values = c("#457b9d","#1d3557"))
#Rip off the legend
  legend <- get_legend(p1)
p1 <- p1 +theme(legend.position="none")

# Build the second figure
p2 <- ggplot(data=df2, aes(x=reorder(Organism, -Cells), y=Cells, fill=Treatment))+
  #ggtitle(~italic("Lactobacillus kunkeei")~"and Acetic Acid Bacteria")+
  theme(axis.title.x = element_blank(),
  legend.position="none",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    text=element_text(family = "sans"),
    panel.background = element_rect(fill='white', color = 'white'))+
  geom_bar(stat="identity", position=position_dodge())+
  labs(tag = 'B')+ #add letters to fulfill multipanel requirements from Catalyst
  scale_fill_manual(values = c("#457b9d","#1d3557"))
p2

# Build the third figure
p3 <- ggplot(data=df3, aes(x=reorder(Organism, -Cells), y=Cells, fill=Treatment))+
  #ggtitle(~italic("Pediococcus")~" species, Zygosaccharomyces, L. brevis/hilgardii/fermentum, and Lactobacillus plantarum/casei/mali")+
  theme(axis.title.x = element_blank(), #no title
  legend.position = "none", #no legend
    plot.title = element_text(hjust = 0.5), #this can be removed but it's a req. for catalyst
    plot.subtitle = element_text(hjust = 0.5), #this can be removed but it's a req. for catalyst
    text=element_text(family = "sans"), #this can be removed but it's a req. for catalyst
    panel.background = element_rect(fill='white', color = 'white'))+ #white background for the plot
  geom_bar(stat="identity", position=position_dodge())+ #bar plot and spacing
  labs(tag = 'C')+ #add letters to fulfill multipanel requirements from Catalyst
  scale_fill_manual(values = c("#457b9d","#1d3557")) #blue colors for plotting
p3

#GROB Layout and Arrange in a Grid
lay <- rbind(c(1,1,1,2,2,2,NA),
             c(1,1,1,2,2,2,NA),
             c(1,1,1,2,2,2,4),
             c(1,1,1,2,2,2,4),
             c(3,3,3,3,3,3,NA),
             c(3,3,3,3,3,3,NA))

#Arrange the images using the layout, add the title, add legend
title1=textGrob("Organismal Response to Treatment")
  theme(element_text(family = "sans"))
grid.arrange(p1,p2,p3,legend, layout_matrix=lay, top=title1)


# Remove the tiff for Catalyst
# dev.off()
```